### 2021-11-03 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@fc8297696fc28950cf3e4169b1fe815ccd7c9ca4 🚀
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@f43705e294b97501d82ade71305b5b18023fb583 🚀

\vfill
Despite their importance in many ecological processes, collecting data
and information on ecological interactions is an exceedingly challenging
task. For this reason, large parts of the world have a data deficit when
it comes to species interactions, and how the resulting networks are
structured. As data collection alone is unlikely to be sufficient,
community ecologists must adopt predictive methods. Here we develop such
a method, relying on graph embedding and transfer learning to assemble a
predicted list of trophic interactions between Canadian mammals. This
interaction list is derived from the European food web, despite sharing
only 5\% of common species with Canada. The results of the predictive
model are compared against databases of recorded pairwise interactions,
showing that we correctly recover over 95\% of known interactions. We
provide guidance on how this method can be adapted by substituting some
approaches or predictors in order to make it more generally applicable.

@@

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen
2016). First, networks within a location are difficult to sample
properly (Jordano 2016a, b), resulting in a widespread ``Eltonian
shortfall'' (Hortal \emph{et al.} 2015), \emph{i.e.} a lack of knowledge
about inter and intra specific relationships. This first challenge has
been, in large part, addressed by the recent emergence of a suite of
methods aiming to predict interactions within \emph{existing} networks,
many of which are reviewed in Strydom \emph{et al.} (2021a). Second,
recent analyses based on collected data (Poisot \emph{et al.} 2021a) or
metadata (Cameron \emph{et al.} 2019) highlight that ecological networks
are currently studied in a biased subset of space and bioclimates, which
impedes our ability to generalize any local understanding of network
structure. Meaning that, although the framework to address
incompleteness \emph{within} networks exists, there would still be
regions for which, due to a \emph{lack} of local interaction data, we
are unable to infer potential species interactions. Having a general
solution for inferring a \emph{plausible} metaweb (despite the
unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales.

Here, we present a general method for the transfer learning of network
representations, relying on the similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. This allows us to
construct a \emph{probabilistic} metaweb for a community for which we
have \emph{no} prior trophic interaction data for the desired species
pool. Our methodology is outlined in fig.~\ref{fig:concept}, where we
provide an illustration based on learning the embedding of a metaweb of
trophic interactions for European mammals (known interactions; Maiorano
\emph{et al.} 2020b, a) and, based on phylogenetic relationships between
mammals globally (\emph{i.e.}, phylogenetic tree Upham \emph{et al.}
2019), infer a metaweb for the Canadian mammalian species pool
(interactions are treated as unknown in this instance).

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (vulnerability traits representing species at
the lower trophic-level and generality traits representing species at
higher trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

There is a plurality of measures of species similarities that can be
used for metaweb reconstruction (see \emph{e.g.} Morales-Castilla
\emph{et al.} 2015); however, phylogenetic proximity has several
desirable properties when working at large scales. Gerhold \emph{et al.}
(2015) made the point that phylogenetic signal captures diversification
of characters (large macro-evolutionary process), but not necessarily
community assembly (fine ecological process); Dormann \emph{et al.}
(2010) previously found very similar conclusions. Interactions tend
reflect a phylogenetic signal because they have a conserved pattern of
evolutionary convergence that encompasses a wide range of ecological and
evolutionary mechanisms (Cavender-Bares \emph{et al.} 2009; Mouquet
\emph{et al.} 2012), and - most importantly - retain this signal even
when it is not detectable at the community scale (Hutchinson \emph{et
al.} 2017; Poisot \& Stouffer 2018). Finally, species interactions at
macro-ecological scales seem to respond mostly to macro-evolutionary
processes (Price 2003); which is evidenced by the presence of conserved
backbones in food webs (Dalla Riva \& Stouffer 2016), strong
evolutionary signature on prey choice (Stouffer \emph{et al.} 2012), and
strong phylogenetic signature in food web intervality (Eklöf \& Stouffer
2016). Phylogenetic reconstruction has also previously been used within
the context of ecological networks, namely understanding ancestral
plant-insect interactions (Braga \emph{et al.} 2021). Taken together,
these considerations suggest that phylogenies can reliably be used to
transfer knowledge on species interactions.

Our case study shows that phylogenetic transfer learning is indeed an
effective approach to predict the Canadian mammalian metaweb. This
showcases that although the components (species) that make up the
Canadian and European communities may be \emph{minimally} shared, if the
medium (proxy space) selected in the transfer step is biologically
plausible, we can still effectively learn from the known network and
make biologically relevant predictions of interactions. It should be
reiterated that the framework presented in fig.~\ref{fig:concept} is
amenable to changes; notably, the measure of similarity may not be
phylogeny, and can be replaced by information on foraging (Beckerman
\emph{et al.} 2006), cell-level mechanisms (Boeckaerts \emph{et al.}
2021), or a combination of traits and phylogenetic structure (Stock
2021).

@@

We use data from the European metaweb assembled by Maiorano \emph{et
al.} (2020b), following the definition of the metaweb first introduced
by Dunne (2006), \emph{i.e.} an inventory of all possible interactions
within species from a spatially delimited pool. Notably the metaweb is
not a prediction of the food web at any specific locale within the
frontiers of the species pool -- in fact, these local food webs are
expected to have a subset of both the species and the interactions of
their metaweb (Poisot \emph{et al.} 2012). This being said, as the
metaweb represents the total of functional, phylogenetic, and
macroecological processes (Morales-Castilla \emph{et al.} 2015), it is
thus still worthy of ecological attention. We deduced the subgraph
corresponding to all mammals by matching species names in the original
network to the GBIF taxonomic backbone (GBIF Secretariat 2021) and
retaining all those who matched to mammals. This serves a dual purpose
1) to extract only mammals from the European network and 2) to match and
standardize species names when aggregating the different data sources
further downstream (which is an important consideration when combining
datasets (Grenié \emph{et al.} 2021)). All nodes had valid matches to
GBIF at this step, and so this backbone is used for all name
reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer. We used the mammalian consensus supertree by
Upham \emph{et al.} (2019), for which all approximatively 6000 names
have been similarly matched to their GBIF valid names. This step allows
us to place each node of the mammalian European metaweb in the
phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by Maiorano \emph{et al.} (2020b) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

The crux of the method is the transfer of knowledge of a known network,
in order to predict interactions between species from another location.
In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko \emph{et al.} 2011) to generate latent traits representing
a low-dimensional embedding of the network; these traits give an
unbiased estimate of the node's position in the latent feature spaces.
Then, we map these latent traits onto a reference phylogeny (other
distance-based measures of species proximity that allow for the
inference of features in the latent space can be used, for example the
dissimilarity in functional traits). Based on the reconstructed latent
traits for species in the destination species pool, a Random Dot Product
Graph model (hereafter RDPG; Young \& Scheinerman 2007) predicts the
interaction between species through a function of the nodes' features
through matrix multiplication. Thus, from latent traits and node
position, we can infer interactions.

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson
\emph{et al.} 2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl} (Dansereau \&
Poisot 2021). The network embedding and analysis is done using
\texttt{EcologicalNetworks.jl} (Poisot \emph{et al.} 2019; Banville
\emph{et al.} 2021). The phylogenetic simulations are done using
\texttt{PhyloNetworks.jl} (Solís-Lemus \emph{et al.} 2017) and
\texttt{Phylo.jl} (Reeve \emph{et al.} 2016). A complete
\texttt{Project.toml} file specifying the full tree of dependencies is
available alongside the code. This material also includes a fully
annotated copy of the entire code required to run this project
(describing both the intent of the code and discussing some technical
implementation details), a vignette for every step of the process, and a
series of Jupyter notebooks with the text and code. The pipeline can be
executed on a laptop in a matter of minutes, and therefore does not
require extensive computational power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016). In addition, recent advances show that the
latent variables produced this way can be used to predict \emph{de novo}
network edges (\emph{i.e.} interactions; Runghen \emph{et al.} 2021).

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom
\emph{et al.} 2021b) and to allow highly accurate prediction of the
interactions within a single network (Poisot \emph{et al.} 2021b). Under
SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{L}\mathbf{\Sigma}\mathbf{R'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{L}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{R}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{L}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{R}'\) a \(n \times r\) semi-unitary matrix.

The specific rank at which the SVD ought to be truncated is a difficult
question. The purpose of SVD is to remove the noise (expressed at high
dimensions) and to focus on the signal, (expressed at low dimensions).
In datasets with a clear signal/noise demarcation, a scree plot of
\(\mathbf{\Sigma}\) can show a sharp drop at the rank where noise starts
(Zhu \& Ghodsi 2006). Because the European metaweb is almost entirely
known, the amount of noise (uncertainty) is low; this is reflected in
fig.~\ref{fig:scree} (left), where the scree plot shows no important
drop, and in fig.~\ref{fig:scree} (right) where the proportion of
variance explained increases smoothly at higher dimensions. For this
reason, we default back to a threshold that explains 60\% of the
variance in the underlying data, corresponding to 12 dimensions -
\emph{i.e.} a tradeoff between accuracy and a reduced number of
features.

A RDPG estimates the probability of observing interactions between nodes
(species) as a function of the nodes' latent variables. The latent
variables used for the RDPG, called the left and right subspaces, are
defined as \(\mathscr{L} = \mathbf{L}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{R}\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R}' = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R}' \approx \mathbf{A}\). Using a rank of 1 for
the t-SVD provides a first-order approximation of the network.

\begin{figure}
\hypertarget{fig:scree}{%
\centering
\includegraphics{figures/figure-screeplot.png}
\caption{Left: representation of the screeplot of the singular values
from the t-SVD on the European metaweb. The screeplot shows no obvious
drop in the singular values that may be leveraged to automatically
detect a minimal dimension for embedding, after \emph{e.g.} Zhu \&
Ghodsi (2006). Right: cumulative fraction of variance explained by each
dimension up to the rank of the European metaweb. The grey lines
represent cutoffs at 50, 60\ldots{} 90\% of variance explained. For the
rest of the analysis, we reverted to an arbitrary threshold of 60\% of
variance explained, which represented a good tradeoff between accuracy
and reduced number of features.}\label{fig:scree}
}
\end{figure}

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano \emph{et al.} (2020b) and O'Connor \emph{et al.} (2020), this
seems like a reasonable assumption, although one that does not hold for
all metawebs. We used the thresholding approach presented in Poisot
\emph{et al.} (2021b), and picked a cutoff that maximized Youden's \(J\)
statistic (a measure of the informedness (trust) of predictions; Youden
(1950)); the resulting cutoff was 0.22, and gave an accuracy above 0.99.

>>>>

\begin{enumerate}
    \itemDespite their importance in many ecological processes,
collecting data and information on ecological interactions is an
exceedingly challenging task. For this reason, large parts of the world
have a data deficit when it comes to species interactions, and how the
resulting networks are structured. As data collection alone is unlikely
to be sufficient, community ecologists must adopt predictive methods.%
    \itemWe present a methodological framework that uses graph embedding
and transfer learning to assemble a predicted list of trophic
interactions of a species pool for which their interactions are unknown.
Specifically, we `learn' the information from a known interaction
network by inferring the latent traits of species and infer the latent
traits of a species pool for which we have no \emph{a priori}
interaction data based on their phylogenetic relatedness to species from
the known network. The latent traits can then be used to predict
interactions and construct and interaction network.%
    \itemHere we assembled a metaweb for Canadian mammals derived from
interactions in the European food web, despite only 4\% of common
species being shared between the two locations. The results of the
predictive model are compared against databases of recorded pairwise
interactions, showing that we correctly recover 91\% of known
interactions.%
    \itemThe framework itself is robust even when the known network is
incomplete or contains spurious interactions making it an ideal
candidate as a tool for filling gaps when it comes to species
interactions. We provide guidance on how this frameworl can be adapted
by substituting some approaches or predictors in order to make it more
generally applicable.%
\end{enumerate}

@@

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen,
2016). First, ecological networks within a location are difficult to
sample properly (Jordano, 2016a, 2016b), resulting in a widespread
``Eltonian shortfall'' (Hortal et al., 2015), \emph{i.e.} a lack of
knowledge about inter- and intra- specific relationships. This first
challenge has been, in large part, addressed by the recent emergence of
a suite of methods aiming to predict interactions within \emph{existing}
networks, many of which are reviewed in Strydom, Catchen, et al. (2021).
Second, recent analyses based on collected data (Poisot, Bergeron, et
al., 2021) or metadata (Cameron et al., 2019) highlight that ecological
networks are currently studied in a biased subset of space and
bioclimates, which impedes our ability to generalize any local
understanding of network structure. Meaning that, although the framework
to address incompleteness \emph{within} networks exists, there would
still be regions for which, due to a \emph{lack} of local interaction
data, we are unable to infer potential species interactions. Having a
general solution for inferring \emph{potential} interactions (despite
the unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom, Catchen, et al.
(2021) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

Here, we present a general method to recommend potential trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang, 2010;
Torrey \& Shavlik, 2010). In this instance, we solve the problem of
predicting trophic interactions between species, based on knowledge
extracted from another species pool for which interactions are known by
using phylogenetic structure as a medium for transfer. There is a
plurality of measures of species similarities that can be used for
inferring \emph{potential} species interactions \emph{i.e.} metaweb
reconstruction (see \emph{e.g.} Morales-Castilla et al., 2015); however,
phylogenetic proximity has several desirable properties when working at
large scales. Gerhold et al. (2015) made the point that phylogenetic
signal captures diversification of characters (large macro-evolutionary
process), but not necessarily community assembly (fine ecological
process); Dormann et al. (2010) previously found very similar
conclusions. Interactions tend to reflect a phylogenetic signal because
they have a conserved pattern of evolutionary convergence that
encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares et al., 2009; Mouquet et al., 2012), and - most
importantly - retain this signal even when it is not detectable at the
community scale (Hutchinson et al., 2017; Poisot \& Stouffer, 2018).
Finally, species interactions at macro-ecological scales seem to respond
mostly to macro-evolutionary processes (Price, 2003); which is evidenced
by the presence of conserved backbones in food webs (Dalla Riva \&
Stouffer, 2016; Mora et al., 2018), strong evolutionary signature on
prey choice (Stouffer et al., 2012), and strong phylogenetic signature
in food web intervality (Eklöf \& Stouffer, 2016). Phylogenetic
reconstruction has also previously been used within the context of
ecological networks, namely understanding ancestral plant-insect
interactions (Braga et al., 2021). Taken together, these considerations
suggest that phylogenies can reliably be used to transfer knowledge on
species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (latent vulnerability traits are more
representative of species at the lower trophic-level and latent
generality traits are more representative of species at higher
trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano et al.,
2020a, 2020b) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham et al., 2019), infer a
metaweb for the Canadian mammalian species pool (interactions are
treated as unknown in this instance). Following the definition of Dunne
(2006), a metaweb is a network analogue to the regional species pool;
specifically, it is an inventory of all \emph{potential} interactions
between species from a spatially delimited area (and so captures the
\(\gamma\) diversity of interactions). The metaweb is, therefore,
\emph{not} a prediction of the food web at a specific locale within the
spatial area it covers, and will have a different structure (notably by
having a larger connectance; see \emph{e.g.} Wood et al., 2015). These
local food webs (which captures the \(\alpha\) diversity of
interactions) are a subset of the metaweb's species and interactions,
and have been called ``metaweb realizations'' (Poisot et al., 2015).
Differences between local food web and their metaweb are due to chance,
species abundance and co-occurrence, local environmental conditions, and
local distribution of functional traits, among others.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it is the
``upper bounds'' on what the composition of the local networks can be
(see \emph{e.g.} McLeod et al., 2021). These local networks, in turn,
can be reconstructed given appropriate knowledge of local species
composition, providing information on structure of food webs at finer
spatial scales. This has been done for example for
tree-galler-parasitoid systems (Gravel et al., 2018), fish trophic
interactions (Albouy et al., 2019), tetrapod trophic interactions
(O'Connor et al., 2020), and crop-pest networks (Grünig et al., 2020).
Whereas the original metaweb definition, and indeed most past uses of
metawebs, was based on the presence/absence of interactions, we focus on
\emph{probabilistic} metawebs where interactions are represented as the
chance of success of a Bernoulli trial (see \emph{e.g.} Poisot et al.,
2016); therefore, not only does our method recommend interactions that
may exist, it gives each interaction a score, allowing us to properly
weigh them.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes; notably, the measure
of similarity may not be phylogeny, and can be replaced by information
on foraging (Beckerman et al., 2006), cell-level mechanisms (Boeckaerts
et al., 2021), or a combination of traits and phylogenetic structure
(Stock, 2021). Most importantly, although we focus on a trophic system,
it is an established fact that different (non-trophic) interactions do
themselves interact with and influence the outcome of trophic
interactions (see \emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012).
Future development of metaweb inference techniques should cover the
prediction of multiple interaction types.

@@

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
This metaweb itself is a network of binary (\emph{i.e.}
presence/absence), potential two-way interactions between species pairs.

We filtered down the European metaweb to create a subgraph corresponding
to all mammals by matching species names in the original network to the
Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF
Secretariat, 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié et al., 2021)). All nodes
had valid matches to GBIF at this step, and so this backbone is used for
all name reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer (\emph{i.e.} the transfer medium). We used the
mammalian consensus supertree by Upham et al. (2019), for which all
approximatively 6000 names have been similarly matched to their GBIF
valid names. This step allows us to place each node of the mammalian
European metaweb in the phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the International Union for Conservation of Nature
(IUCN) checklist, and selected the terrestrial and semi-aquatic species
(this corresponds to the same selection that was applied by Maiorano et
al. (2020a) in the European metaweb). The IUCN names were, as
previously, reconciled against GBIF to have an exact match to the
taxonomy.

@@

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location at which the network is unknown (or partially
known). In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko et al., 2011) to generate latent traits representing a
low-dimensional embedding of the network. As an aside, most ecologists
are indirectly familiar with SVD: Principal Component Analysis is a
special case of SVD, which is more sensitive to numerical instabilities
(see notably Shlens, 2014). The latent traits give an unbiased estimate
of the node's position in the latent feature spaces and can be mapped
onto a reference phylogeny (other distance-based measures of species
proximity that allow for the inference of features in the latent space
can be used, for example the dissimilarity in functional traits). Based
on the reconstructed latent traits for species in the destination
species pool, a Random Dot Product Graph model (hereafter RDPG; S. J.
Young \& Scheinerman, 2007) predicts the interaction between species
through a function of the nodes' features through matrix multiplication.
Thus, from latent traits and node position, we can infer interactions.

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm, 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy et al., 2009); in fact, in an instance of ecological forecasting
of spatio-temporal systems, model-free approaches (\emph{i.e.} drawing
all of their information from the data) outperformed model-informed ones
(Perretti et al., 2013).

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson et al.,
2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl} (Dansereau \&
Poisot, 2021). The network embedding and analysis is done using
\texttt{EcologicalNetworks.jl} (Banville et al., 2021; Poisot et al.,
2019). The phylogenetic simulations are done using
\texttt{PhyloNetworks.jl} (Solís-Lemus et al., 2017) and
\texttt{Phylo.jl} (Reeve et al., 2016). A complete \texttt{Project.toml}
file specifying the full tree of dependencies is available alongside the
code. This material also includes a fully annotated copy of the entire
code required to run this project (describing both the intent of the
code and discussing some technical implementation details), a vignette
for every step of the process, and a series of Jupyter notebooks with
the text and code. The pipeline can be executed on a laptop in a matter
of minutes, and therefore does not require extensive computational
power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer, 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically similar
predators should share phylogenetically similar preys. In addition,
recent advances show that the latent variables produced this way can be
used to predict \emph{de novo} network edges. Interestingly, the latent
variables do not need to be produced by decomposing the network itself;
in a recent contribution, Runghen et al. (2021) showed that deep
artificial neural networks are able to reconstruct the left and right
subspaces of an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional metadata as predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom, Dalla
Riva, et al., 2021) and to allow highly accurate prediction of the
interactions within a single network (Poisot, Ouellet, et al., 2021).
Under SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.

The specific rank at which the SVD ought to be truncated is a difficult
question. The purpose of SVD is to remove the noise (expressed at high
dimensions) and to focus on the signal, (expressed at low dimensions).
In datasets with a clear signal/noise demarcation, a scree plot of
\(\mathbf{\Sigma}\) can show a sharp drop at the rank where noise starts
(Zhu \& Ghodsi, 2006). Because the European metaweb is almost entirely
known, the amount of noise (uncertainty) is low; this is reflected in
fig.~\ref{fig:scree} (left), where the scree plot shows no important
drop, and in fig.~\ref{fig:scree} (right) where the proportion of
variance explained increases smoothly at higher dimensions. For this
reason, we default back to a threshold that explains 60\% of the
variance in the underlying data, corresponding to 12 dimensions -
\emph{i.e.} a tradeoff between accuracy and a reduced number of
features.

An RDPG estimates the probability of observing interactions between
nodes (species) as a function of the nodes' latent variables, and is a
way to turn a SVD (which decompose one matrix into three) into two
matrices that can be multiplied to provide an approximation of the
network. The latent variables used for the RDPG, called the left and
right subspaces, are defined as
\(\mathscr{L} = \mathbf{U}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{V}'\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R} = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R} \approx \mathbf{A}\). Using a rank of 1 for the
t-SVD provides a first-order approximation of the network. One advantage
of using a RDPG rather than a SVD is that the number of components to
estimate decreases; notably, one does not have to estimate the singular
values of the SVD. Furthermore, the two subspaces can be directly
multiplied to yield a network.

\begin{figure}
\hypertarget{fig:scree}{%
\centering
\includegraphics{figures/figure-screeplot.png}
\caption{Left: representation of the scree plot of the singular values
from the t-SVD on the European metaweb. The scree plot shows no obvious
drop in the singular values that may be leveraged to automatically
detect a minimal dimension for embedding, after \emph{e.g.} Zhu \&
Ghodsi (2006). Right: cumulative fraction of variance explained by each
dimension up to the rank of the European metaweb. The grey lines
represent cutoffs at 50, 60, \ldots, 90\% of variance explained. For the
rest of the analysis, we reverted to an arbitrary threshold of 60\% of
variance explained, which represented a good tradeoff between accuracy
and reduced number of features.}\label{fig:scree}
}
\end{figure}
